Batched Bayesian Optimization via Multi-objective Acquisition Ensemble for Automated Analog Circuit Design

### 2018-01-30 paper/abstract.tex
Paper update

\begin{abstract} 
    The Bayesian optimization has recently shown promising performance for
    automated analog circuit design, however, the sequential property of
    Bayesian optimization limits its further application. In this paper, a
    batched bayesian optimization method is proposed, the parallelization is
    realized via a multi-objective ensemble of acquisition functions. In each
    iteration, multiple acquisition functions are selected, the multi-objective
    optimization is then performed to search for the Pareto front of the
    acquisition functions, candidate points are then sampled from the Pareto
    fronts. The proposed method is compared with several state-of-the-art
    batched Bayesian optimization algorithms using analytical benchmark
    functions and real-world analog circuits; the experimental results show
    that the proposed method is competitive with state-of-the-art methods.
\end{abstract}

>>>>

\begin{abstract} 
    The Bayesian optimization has recently shown promising performance for
    automated analog circuit design. However, the sequential property of
    Bayesian optimization limits its further application. In this paper, a
    batched bayesian optimization method is proposed, the parallelization is
    realized via a multi-objective ensemble of acquisition functions. In each
    iteration, multiple acquisition functions are selected, the multi-objective
    optimization is then performed to search for the Pareto front of the
    acquisition functions, candidate points are then sampled from the Pareto
    fronts. The proposed method is compared with several state-of-the-art
    batched Bayesian optimization algorithms using analytical benchmark
    functions and real-world analog circuits; the experimental results show
    that the proposed method is competitive compared to the state-of-the-art
    methods.
\end{abstract}

### 2018-01-30 paper/background.tex
Paper update

We model the $\mathrm{FOM}(\bm{x})$ in \eqref{eq:Formulation} with Gaussian
process (GP) model~\cite{GPML}. The GP model is the most commonly used model
for Bayesian optimization, the advantage of GP is that it can give a
well-calibrated uncertainty of prediction. GP is characterized by a mean
function $m(\bm{x})$ and a covariance function $k(\bm{x}, \bm{x'})$. In this
paper, we use squared-exponential ARD kernel, and a constant mean fucntion
$m(\bm{x}) = \mu_0$ for all our experiments.

@@

In Bayesian optimization described in Algorithm~\ref{alg:BayesianOptAlgo}, the acquisition function is used to balance the exploration and exploitation during the optimization. The acquisition fucntion considers both the predictive value and the uncertainty. There are a lot of existing acquisition functions, examples include the lower confidence bound (LCB), expected improvement (EI) and probability of improvement (PI).

The LCB fucntion is defined as follows:
\begin{equation}
    \label{eq:LCB}
\end{equation}
Where the $\mu(\bm{x})$ and the $\sigma(\bm{x})$ are the predictive value and uncertainty of GP. $\kappa$ is a parameter that balances the exploitation and exploration.

@@

>>>>

We model the $\mathrm{FOM}(\bm{x})$ in \eqref{eq:Formulation} with Gaussian
process (GP) model~\cite{GPML}. The GP model is the most commonly used model
for Bayesian optimization, the advantage of GP is that it can give a
well-calibrated uncertainty of prediction. GP is characterized by a mean
function $m(\bm{x})$ and a covariance function $k(\bm{x}, \bm{x'})$. In this
paper, we use squared-exponential ARD kernel, and a constant mean function
$m(\bm{x}) = \mu_0$ for all our experiments.

@@

In Bayesian optimization described in Algorithm~\ref{alg:BayesianOptAlgo}, the acquisition function is used to balance the exploration and exploitation during the optimization. The acquisition function considers both the predictive value and the uncertainty. There are a lot of existing acquisition functions, examples include the lower confidence bound (LCB), the probability of improvement (PI), and the expected improvement (EI).

The LCB function is defined as follows:
\begin{equation}
    \label{eq:LCB}
    \mathrm{LCB}(\bm{x}) = \mu(\bm{x}) - \kappa \sigma(\bm{x})
\end{equation}
Where the $\mu(\bm{x})$ and the $\sigma(\bm{x})$ are the predictive value and uncertainty of GP. $\kappa$ is a parameter that balances the exploitation and exploration.

@@

The PI and EI function are defined as
\begin{equation}
    \label{eq:PI_EI}
    \begin{array}{lll}
        \mathrm{PI}(\bm{x}) &=& \Phi(\lambda) \\
        \mathrm{EI}(\bm{x}) &=& \sigma(\bm{x}) (\lambda \Phi(\lambda) + \phi(\lambda))     \\
        \mathrm{\lambda}    &=& \displaystyle \frac{\tau - \xi - \mu(\bm{x})}{\sigma(\bm{x})}, 
    \end{array}
\end{equation}
where $\tau$ is the current best value objective value, and $\xi$ is a small positive jitter to improvement the ability of exploration. The $\Phi(.)$ and $\phi(.)$ functions are the CDF and PDF functions of normal distribution.

% TODO: discuss the different behaviours of LCB, EI, PI

There are also other acquisition function, like the knowledge gradient~\cite{scott2011correlated} function, predictive entropy search~\cite{hernandez2014predictive}. A portforlio of several acquisition function is also possible~\cite{hoffman2011portfolio}.

### 2018-01-30 paper/intro.tex
Paper update

% TODO: Traditional methods using offline model and simulated based methods are
% not good
The analog circuit design automation problems can be formulated as optimization
problems, the aim is to find the optimal vector of design parameters that give
best figure of merit (FOM). Prior works about analog circuit optimization
include offline model-based approaches
~\cite{colleran2003optimization,daems2003simulation,wang2014enabling} and
simulation based approaches. The offline model-based methods try to build
accurate global model and then apply global optimization algorithms to the
cheap-to-evaluate models, the model either comes from designers' manual
derivation or from regression models like SVM and ANN. The problem of this
approach is that the accurate models are usually hard to get, for example,
in~\cite{wang2014enabling}, 100,000 randomly simulated points are used to train
a sparse polynomial model for a circuit with ten design parameters.
Simulation-based methods, instead, treat the circuits as black-box functions
that get the objective value from circuit simulations, global optimization
algorithms are directly applied to the black-box functions. For
simulation-based circuit optimization methods, meta-heuristic algorithms like
particle swarm intelligence (PSO)~\cite{phelps2000anaconda} and differential
evolution algorithm (DE)~\cite{liu2009analog} are widely used, although these
algorithm are able to explore the whole design space, they have relatively low
convergence rate and consume many circuit simulations. When the circuit
simulation takes a long time, both model-based and simulation-based approaches
can be prohibitive.

% TODO: Bayesian optimization is a sequential algorithm, there is a need to
% parallelize it
%
% TODO: cite the two 3D-IC papers of TVLSI
To reduce the number of circuit simulations, the Gaussian process
(GP)~\cite{GPML} model has been introduced as online surrogate model to assist
the optimization. In~\cite{liu2014gaspad}, GP is combined with differential
evolution algorithm, in~\cite{lyu2017efficient}. Recently, the Bayesian
optimization (BO)~\cite{shahriari2016taking} algorithm has been applied for
analog circuit optimization, the Bayesian optimization algorithm is firstly
introduced for the automated design of general analog circuits, and has shown
very good performance compared to other simulation-based approaches,
in~\cite{wang2017efficient}, the Bayesian optimization algorithm is combined
with adaptive Monte-Carlo sampling to optimize the yield of analog circuits and
static random-access memory (SRAM).

The Bayesian optimization algorithm is a well-studied algorithm, and has
demonstrated to be a promising algorithm for automated analog circuit design,
however, the standard Bayesian optimization algorithm is sequential, it chooses
the next evaluation point by optimizing the specified acquisition fucntion,
which makes the parallelization of Bayesian optimization non-trivial. The
sequential property of BO limits its further application, as multi-core
processors are usually availble for modern computers.

@@

% TODO: Beiefly introduce my algorithm
All the above mentioned algorithms choose to use one acquisition function, and
except for the smulation mathcing method and local penalization method which
can select arbitrary acquisition function, other parallelization methods rely on
specific acquisition function, the UCB acquisition fucntion must be used for
BUCB and GP-UCB-PE, and the knowledge gradient method must be used for qKG
algorithm. As has been stated in~\cite{hoffman2011portfolio}, no one
acquisition fucntion can always outperform other acquisition functions, so
relying on one acquisition fucntion may resulting in poor performance. Also, a
portfolio strategy that ensembles multiple acquisition functions would be
helpful.

In this paper, we propose to parallize the Bayesian optimization algorithm via
a multi-objective ensemble of acquisition functions. Firstly, in each
iteration, after the GP model is updated, multiple acquisition functions are
selected, we then perform multi-objective optimization to find the \emph{Pareto
front} (PF) of the acquisition functions. The PF represents the best trade-off
between these acquisition functions, when used in sequential mode, the propose
multi-objective acquisition ensemble (MACE) strategy can be seen as a portfolio
strategy, when batched evaluation is possible, we can sample multiple points on
the PF as there are usually much more points on the PF than the batch size.

We tested the MACE algorithm using several analytical benchmark functions and
two real-world analog circuits, including an operational amplifier with
\textcolor{red}{14} design parameters and a class-E power amplifier with
\textcolor{red}{ten} design parameters. The BLCB method, local penalization
method with expected improvement acquisition fucntion (EI-LP), the qEI methods
and qEI methods are compared with MACE. The proposed MACE method achieved
competitive performance when compared with the listed state-of-the-art
algorithms.

>>>>

% TODO: Traditional methods using offline model and simulated based methods are
% not good
The analog circuit design automation problems can be formulated as optimization
problems, the aim is to find the optimal vector of design parameters that give
the best figure of merit (FOM). Prior works about analog circuit optimization
include offline model-based approaches
~\cite{colleran2003optimization,daems2003simulation,wang2014enabling} and
simulation-based approaches. The offline model-based methods try to build
accurate global models and then apply global optimization algorithms to the
cheap-to-evaluate models. The model either comes from designers' manual
derivation or regression models like SVM and ANN. The problem of this approach
is that the accurate models are usually hard to get, for example,
in~\cite{wang2014enabling}, 100,000 randomly simulated points are used to train
a sparse polynomial model for a circuit with ten design parameters.
Simulation-based methods, instead, treat the circuits as black-box functions
that get the FOM value from circuit simulations, global optimization algorithms
are then directly applied to the black-box functions. For simulation-based
circuit optimization methods, meta-heuristic
algorithms~\cite{phelps2000anaconda, liu2009analog} are widely used, although
these algorithms can explore the whole design space, they have relatively low
convergence rate and consume many circuit simulations. When the circuit
simulation takes a long time, both model-based and simulation-based approaches
can be very time-consuming.

% TODO: Bayesian optimization is a sequential algorithm, there is a need to
% parallelize it
%
% TODO: cite the two 3D-IC papers of TVLSI
To reduce the number of circuit simulations, the Gaussian process
(GP)~\cite{GPML} model has been introduced as an online surrogate model to
assist the optimization. In~\cite{liu2014gaspad}, GP is combined with
differential evolution algorithm, Recently, the
Bayesian optimization (BO)~\cite{shahriari2016taking} algorithm has been
applied for analog circuit optimization. In~\cite{lyu2017efficient}, the Bayesian optimization algorithm is
firstly introduced for the automated design of general analog circuits and has
shown to be more sample-efficient compared to other simulation-based
approaches. In~\cite{wang2017efficient}, the Bayesian optimization algorithm
is combined with adaptive Monte-Carlo sampling to optimize the yield of analog
circuits and static random-access memory (SRAM).

The Bayesian optimization algorithm is a well-studied algorithm and has
demonstrated to be a promising algorithm for automated analog circuit design.
However, the standard Bayesian optimization algorithm is sequential; it chooses
the next evaluation point by optimizing the specified acquisition function,
which makes the parallelization of Bayesian optimization non-trivial. The
sequential property of BO limits its further application in multi-core computer
systems.

@@

% TODO: Beiefly introduce my algorithm
All the algorithms mentioned above choose to use one acquisition function, and
except for the simulation matching method and local penalization method which
can select arbitrary acquisition function, other parallelization methods rely
on specific acquisition function, the UCB acquisition function must be used for
BUCB and GP-UCB-PE, and the knowledge gradient method must be used for the qKG
algorithm. As has been stated in~\cite{hoffman2011portfolio}, no one
acquisition function can always outperform other acquisition functions, so
relying on one acquisition function may resulting in poor performance.

In contrast to existing batched bayesian optimization methods, we propose a
different perspective to the parallelization of Bayesian optimization. In this
paper, we propose to parallelize the Bayesian optimization algorithm via a
multi-objective ensemble of acquisition functions. Firstly, in each iteration,
after the GP model is updated, multiple acquisition functions are selected, we
then perform multi-objective optimization to find the \emph{Pareto front} (PF)
of the acquisition functions. The PF represents the best trade-off between
these acquisition functions. When used in sequential mode, the proposed
multi-objective acquisition ensemble (MACE) strategy can be seen as a portfolio
strategy, when batched evaluation is possible, we can sample multiple points on
the PF as there are usually much more points on the PF than the batch size.

The MACE algorithm is tested using several analytical benchmark functions and
two real-world analog circuits, including an operational amplifier with
\textcolor{red}{14} design parameters and a class-E power amplifier with
\textcolor{red}{ten} design parameters. The BLCB method, local penalization
method with expected improvement acquisition function (EI-LP), the qEI methods
and qEI methods are compared with MACE. The proposed MACE method achieved
competitive performance when compared with the listed state-of-the-art
algorithms.

### 2018-01-30 paper/mace.tex
Paper update

\subsection{Multi-objective optimization}

@@

\textcolor{red}{Plot a figure to illustrate the PF/PS of acquisition functions}

>>>>

\subsection{Multi-objective Optimization of Acquisition Functions}\label{sec:MOForumlation}

\textcolor{red}{See how other MOBO paper formulate multi-objective optimization}

\textcolor{red}{ALERT: COPIED FROM MY TCAS-I paper}

Multi-objective optimization~\cite{MO_overview} can be formulated as:
\begin{equation}
    \label{eq:MOFormulation}
    \begin{aligned}
        & \text{minimize} & & f_1(\bm{x}),~\dots~,f_m(\bm{x})
    \end{aligned}
\end{equation}

For multi-objective optimization, there are more than one objecive to optimize, and usually there does not exist a single solution that simultaneously optimize all these objectives. A design $\bm{x}_1$ is said to be \emph{dominate} by $\bm{x}_2$ if $\forall i \in \{1\dots m\},~f_i(\bm{x}_1) \le f_i(\bm{x}_2)$ and $\exists j \in \{1\dots m\}, f_j(\bm{x}_1) < f_j(\bm{x}_2)$. A design is \emph{Pareto-optimal} if it is not dominated by any other point and dominate at least one point. The whole set of the non-dominated points in the design space is called the \emph{Pareto set}, and the set of non-dominated points in the objective space is called the \emph{Pareto front}. It is often unlikely to get the whole Pareto front as there might be infinite points on the front, the goal of multi-objective optimization is to find a set of designs that approximates the true Pareto front.

There exist many mature multi-objective optimization algorithms, like the non-dominated sorting based genetic algorithm (NSGA-II)~\cite{nsgaii}, and the multi-objective evolutionary algorithm based on decomposition (MOEA/D)~\cite{moead}. In this paper, the multi-objective optimization based on differential evolution (DEMO)~\cite{demo} is used to solve multi-objective optimization problems.

% When we have more than one objective to optimize, the problem is formulated as:

@@

We propose a novel heuristic for the parallelization of Bayesian optimization.
The parallelization is realized via multi-objective ensemble of multiple
acquisition functions. When we have multiple acquisition functions built from
the same GP model, they may not disagree with each other about which point is
most promising, for example, the value of LCB function always decreases as the
$\sigma(\bm{x})$ increases, however, for the PI function, when $\sigma{\bm{x}}$
increases, the value of PI would decrease when $\mu(\bm{x}) < \tau$, and
increase when $\mu(\bm{x}) > \tau$.

% TODO: discuss the difference of acquisition functions: LCB: do not avoid repeated sampling, EI: always positive

With multi-objective optimization, the disagreement between different
acquisition function can be used for the parallelization of Bayesian
optimization. In the standard Bayesian optimization, one acquisition function
is specified, and in each iteration, the single-objective optimization is
performed to optimize the acquisition function; in our proposed MACE algorithm,
multiple acquisition function is pre-specified by the user, in each iteration,
the DEMO multi-objective optimization algorithm is applied to find the Pareto
front of the acquisition functions. Once the Pareto front is obtained, we can
then sample from the Pareto front, and then evaluate these sampled points in
parallel.

% TODO: We illustrate the proposed MACE algorithm using the Branin-Hoo function~\cite{dixon1978global}
% TODO: plot contour of Branin/Ei/LCB/PI, plot the PS and the PF

We select the LCB, EI and PI acquisition function in our implementation of the MACE algorithm, but other acquisition functions like KG and PES can also be incorporated into the MACE framework. The proposed MACE algorithm is described in Algorithm~\ref{alg:MACE}.

\begin{algorithm}
\caption{Multi-objective Acquisition Ensemble Algorithm}
\label{alg:MACE}
\begin{algorithmic}[1]
\STATE Initial Sampling
\STATE Construct initial GP model
\FOR{t = 1, 2, \dots}
    \STATE Construct the LCB, EI and PI function according to \eqref{eq:LCB} and \eqref{eq:PI_EI}
    \STATE Find the Pareto front of LCB, EI, PI function using the DEMO algorithm
    \STATE Randomly sample B points $\bm{x}_1, \dots, \bm{x}_B$ from the Pareto front
    \STATE Evaluate $\bm{x}_1, \dots, \bm{x}_B$ to get $y_1 = f(\bm{x}_1),~\dots~,y_B = f(\bm{x}_B)$
    \STATE Update the GP model
\ENDFOR
\STATE Return best $f(\bm{x})$ recorded during iterations
\end{algorithmic}
\end{algorithm}

### 2018-01-30 paper/main.tex
Paper update

@@

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Submission and Formatting Instructions for ICML 2018}

@@

\twocolumn[
\icmltitle{Batched Bayesian Optimization via Multi-objective Acquisition Function Ensemble for Automated Analog Circuit Design}

>>>>

% Added by wenlong lyu
\usepackage{amsmath}
\usepackage[plain]{fancyref}
\usepackage{bookmark}
\usepackage{bm}
\usepackage{todonotes}

@@

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Batched Bayesian Optimization via Multi-objective Acquisition Ensemble for Automated Analog Circuit Design}

@@

\twocolumn[
\icmltitle{Batched Bayesian Optimization via Multi-objective Acquisition Ensemble for Automated Analog Circuit Design}
